---
title: "Chords analysis with the chorrrds package"
description: |
  A short description of the post.
author:
  - name: Unknown
date: 08-06-2018
output:
  radix::radix_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Intro

`chorrrds` is a package to retrieve and analyse music. It scrapes 
the [*Cifraclub*]() website to download and organize music chords.

The main reason to create `chorrrds` was Bruna's undergrad thesis. In 
her work, she did an end-to-end analysis, exploring feature engineering
techniques to describe and predict musical genres from music chord 
representation.

`chorrrds` can be considered a package for MIR (Music Information 
Retrieval). MIR is a broad area of computational music which extracts 
and processes from unstructured data like sound waves to structured 
data like sheet music or chords.

In this post we'll describe `chorrrds` functions and show 
some examples. Stay tuned!

## Installation

You can install `chorrrds` from your favourite CRAN mirror, simply running:

```{r eval=FALSE}
install.packages("chorrrds")
```

You can also install the latest versions of `chorrrds` from the R-Music GitHub Organization with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("r-music/chorrrds")
```

## Functions

The main function of the package is called `get_chords()` and it
extracts music chords from a specific artist. This happens in 
two steps:
  1. Extraction of song links for each music of an artist with `get_songs`.
  2. Extraction of music chords using the links with `get_chords`.

```{r, warning=FALSE}
library(magrittr)

# Getting the chords for some Janis Joplin songs
songs <- "janis-joplin" %>% 
  chorrrds::get_songs() %>% 
  dplyr::slice(1:10)

chords <- songs %>% 
  dplyr::pull(url) %>% 
  purrr::map(chorrrds::get_chords) %>%
  purrr::map_dfr(dplyr::mutate_if, is.factor, as.character) 

head(chords, 15)
```



## Data

There are many datasets to consider

## Use case

```{r}
chorrrds::search_data("buarque")
songs <- chorrrds::get_songs("buarque")
chords <- chorrrds::get_chords(songs$url[2])
chords_clean <- chorrrds::clean(chords)
```



