---
title: "chorrrds tutorial"
description: |
  A short description of the post.
author:
  - name: jtrecenti
  - name: brunaw
date: 08-05-2018
output:
  radix::radix_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Intro

`chorrrds` is a package to retrieve and analyse music. It scrapes the [*Cifraclub*]() website to download and organize music chords.

The main reason to create `chorrrds` was Bruna's undergrad thesis. In her work, she did an end-to-end analysis, exploring feature engineering techniques to describe and predict musical genres from music chord representation.

`chorrrds` can be considered a package for MIR (Music Information Retrieval). MIR is a broad area of computational music which extracts and processes from unstructured data like sound waves to structured data like sheet music or chords.

In this post we'll describe `chorrrds` main functions and show some examples. Stay tuned!

## Installation

You can install `chorrrds` from your favourite CRAN mirror running

```{r eval=FALSE}
install.packages("chorrrds")
```

You can install the latest versions of chorrrds from the R-Music GitHub Organization with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("r-music/chorrrds")
```

## Functions

## Data

There are many datasets to consider

## Use case

```{r}
chorrrds::search_data("buarque")
songs <- chorrrds::get_songs("buarque")
chords <- chorrrds::get_chords(songs$url[2])
chords_clean <- chorrrds::clean(chords)
```

