---
title: "Parque Dom Pedro II"
output: 
  flexdashboard::flex_dashboard
---

```{r, include=FALSE}
source("functions.R")
source("preamble.R")
```

Temperatura
=====================================  

Column {data-width=500}
-------------------------------------
    
### Ridges plot
    
```{r}
ridges_plot(df_tarde_DPII, month, vars(tp, morning_tp), 
            xlab = "Celsius", ylab = "Month",
            facets_names = c(tp = "Afternoon temperature",
                             morning_tp = "Morning temperature"))
```
   
Column {data-width=500} {.tabset}
-------------------------------------

### Gráfico das séries

```{r}
series_plot(df_tarde_DPII, date, vars(tp, morning_tp, o3_mass_conc),
            xlab = "Year", ylab = "Value",
            facets_names = c(tp = "Afternoon temperature",
                             morning_tp = "Morning temperature",
                             o3_mass_conc = "Ozone"))
```

### Gráfico de dispersão

```{r}
scatter_plot(df_tarde_DPII, o3_mass_conc, vars(tp, morning_tp),
            xlab = "Celsius", ylab = "Ozone",
            facets_names = c(tp = "Afternoon temperature",
                             morning_tp = "Morning temperature"))
```

Radiação
===================================== 

Column {data-width=500}
-------------------------------------
    
### Ridges plot
    
```{r}
ridges_plot(df_tarde_DPII, month, vars(rd, morning_rd), 
            xlab = expression(W/m^2), ylab = "Month",
            facets_names = c(rd = "Afternoon radiation",
                             morning_rd = "Morning radiation"))
```
   
Column {data-width=500} {.tabset}
-------------------------------------

### Gráfico das séries

```{r}
series_plot(df_tarde_DPII, date, vars(rd, morning_rd, o3_mass_conc),
            xlab = "Year", ylab = "Value",
            facets_names = c(rd = "Afternoon radiation",
                             morning_rd = "Morning radiation",
                             o3_mass_conc = "Ozone"))
```

### Gráfico de dispersão

```{r}
scatter_plot(df_tarde_DPII, o3_mass_conc, vars(rd, morning_rd),
            xlab = expression(W/m^2), ylab = "Ozone",
            facets_names = c(rd = "Afternoon radiaiton",
                             morning_rd = "Morning radiation"))
```

Precipitação
=====================================  

Column {data-width=500}
-------------------------------------
    
### Ridges plot
    
```{r}
ridges_plot(df_tarde_DPII, month, vars(pp, morning_pp), 
            xlab = "mm/h", ylab = "Month",
            facets_names = c(pp = "Afternoon precipitation",
                             morning_pp = "Morning precipitation"))
```
   
Column {data-width=500} {.tabset}
-------------------------------------

### Gráfico das séries

```{r}
series_plot(df_tarde_DPII, date, vars(pp, morning_pp, o3_mass_conc),
            xlab = "Year", ylab = "Value",
            facets_names = c(pp = "Afternoon precipitation",
                             morning_pp = "Morning precipitation",
                             o3_mass_conc = "Ozone"))
```

### Gráfico de dispersão

```{r}
scatter_plot(df_tarde_DPII, o3_mass_conc, vars(pp, morning_pp),
            xlab = "mm/h", ylab = "Ozone",
            facets_names = c(pp = "Afternoon precipitation",
                             morning_pp = "Morning precipitation"))
```


Precipitação (categ)
=====================================

Column {data-width=500}
-------------------------------------

### Ridges plot

```{r}
df_tarde_DPII %>% 
  mutate(pp_c = ifelse(pp == 0, "Não", "Sim"),
         morning_pp_c = ifelse(morning_pp == 0, "Não", "Sim")) %>%
  na.omit %>% 
  tidyr::gather(period, pp, pp_c, morning_pp_c) %>% 
  ggplot(aes(x = o3_mass_conc, y = pp, fill = pp)) +
  geom_density_ridges(show.legend = FALSE) +
  facet_wrap(~period, scales = "free",
             labeller = as_labeller(c(pp_c = "Afternoon precipitation",
                                      morning_pp_c = "Morning precipitation"))) +
  theme_bw() +
  labs(y = "Precipitation", x = "Ozone") 
```

Column {data-width=500}
-------------------------------------

### Boxplots

```{r}
df_tarde_DPII %>% 
  mutate(pp_c = ifelse(pp == 0, "Não", "Sim"),
         morning_pp_c = ifelse(morning_pp == 0, "Não", "Sim")) %>%
  na.omit %>% 
  select(date, o3_mass_conc, pp_c, morning_pp_c) %>% 
  na.omit %>% 
  tidyr::gather(period, precipitation, pp_c, morning_pp_c) %>% 
  ggplot(aes(x = period, y = o3_mass_conc, fill = precipitation)) +
  geom_boxplot() +
  theme_bw() +
  labs(fill = "Precipitation", y = "Ozone", x = "Period") +
  scale_x_discrete(labels = c("Morning", "Afternoon"))
```


Umidade
===================================== 

Column {data-width=500}
-------------------------------------
    
### Ridges plot
    
```{r}
ridges_plot(df_tarde_DPII, month, vars(hm, morning_hm), 
            xlab = "%", ylab = "Month",
            facets_names = c(hm = "Afternoon humidity",
                             morning_hm = "Morning humidity"))
```
   
Column {data-width=500} {.tabset}
-------------------------------------

### Gráfico das séries

```{r}
series_plot(df_tarde_DPII, date, vars(hm, morning_hm, o3_mass_conc),
            xlab = "Year", ylab = "Value",
            facets_names = c(hm = "Afternoon humidity",
                             morning_hm = "Morning humidity",
                             o3_mass_conc = "Ozone"))
```

### Gráfico de dispersão

```{r}
scatter_plot(df_tarde_DPII, o3_mass_conc, vars(hm, morning_hm),
            xlab = "%", ylab = "Ozone",
            facets_names = c(hm = "Afternoon humidity",
                             morning_hm = "Morning humidity"))
```

Velocidade do vento
=====================================

Column {data-width=500}
-------------------------------------
    
### Ridges plot
    
```{r}
ridges_plot(df_tarde_DPII, month, vars(ws, morning_ws), 
            xlab = "m/s", ylab = "Month",
            facets_names = c(ws = "Afternoon wind speed",
                             morning_ws = "Morning wind speed"))
```
   
Column {data-width=500} {.tabset}
-------------------------------------

### Gráfico das séries

```{r}
series_plot(df_tarde_DPII, date, vars(ws, morning_ws, o3_mass_conc),
            xlab = "Year", ylab = "Value",
            facets_names = c(ws = "Afternoon wind speed",
                             morning_ws = "Morning wind speed",
                             o3_mass_conc = "Ozone"))
```

### Gráfico de dispersão

```{r}
scatter_plot(df_tarde_DPII, o3_mass_conc, vars(ws, morning_ws),
            xlab = "m/s", ylab = "Ozone",
            facets_names = c(ws = "Afternoon wind speed",
                             morning_ws = "Morning wind speed"))
```

Inversão térmica
=====================================

Column {data-width=500}
-------------------------------------

### Ridges plot

```{r}
df_tarde_DPII %>% 
  na.omit %>% 
  tidyr::gather(period, inversao, dv_ti_0to199m_9am, dv_ti_200to499m_9am) %>%
  mutate(inversao = ifelse(inversao == 1, "Sim", "Não")) %>% 
  ggplot(aes(x = o3_mass_conc, y = inversao, fill = inversao)) +
  geom_density_ridges(show.legend = FALSE) +
  facet_wrap(~period, scales = "free",
             labeller = as_labeller(c(dv_ti_0to199m_9am = "0 a 199 metros",
                                      dv_ti_200to499m_9am = "200 a 499 metros"))) +
  theme_bw() +
  labs(y = "Precipitation", x = "Ozone") 
```

Column {data-width=500}
-------------------------------------

### Boxplots

```{r}
df_tarde_DPII %>%
  select(date, o3_mass_conc, dv_ti_0to199m_9am, dv_ti_200to499m_9am) %>% 
  na.omit %>% 
  tidyr::gather(period, inversao, dv_ti_0to199m_9am, dv_ti_200to499m_9am) %>%
  mutate(inversao = as.factor(inversao)) %>% 
  ggplot(aes(x = period, y = o3_mass_conc, fill = inversao)) +
  geom_boxplot() +
  theme_bw() +
  labs(fill = "Thermal inversion", y = "Ozone", x = "Period") +
  scale_x_discrete(labels = c("0 a 199 metros", "200 a 499 metros")) +
  scale_fill_discrete(labels = c("No", "Yes"))
```

