---
title: "Introduction to the chorrrds package"
description: |
  Music chords analysis!
author:
  - name: "Bruna Wundervald"
    url: https://brunaw.com
    affiliation: Maynooth University
    affiliation_url: https://www.maynoothuniversity.ie/
  - name: "Julio Trecenti"
    url: https://github.com/jtrecenti
    affiliation: Curso-R
    affiliation_url: http://curso-r.com/
date: 08-19-2018
output:
  radix::radix_article:
    self_contained: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Note: This post was update on 17/03/2020 due to some changes with the new version of the package.*

## Intro

`chorrrds` is a package to retrieve and analyse music chords data. 
It scrapes the [*Cifraclub*](cifraclub.com.br/)
website to download and organize music chords.

The main reason to create `chorrrds` was my undergrad thesis. In 
my work, I did an end-to-end analysis, exploring feature engineering
techniques to describe and predict musical genres from music chord 
representation.

`chorrrds` can be considered a package for MIR (Music Information 
Retrieval). MIR is a broad area of computational music which extracts 
and processes music data, from the unstructured ones, as sound waves,
to structured, as sheet music or chords.

In this post we'll describe `chorrrds` functions and show 
some examples. Stay tuned!

## Installation

You can install `chorrrds` from your favourite CRAN mirror, simply running:

```{r, eval = FALSE}
install.packages("chorrrds")
```

You can also install the latest versios of `chorrrds` from the 
R-Music GitHub organization with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("r-music/chorrrds")
```

## Functions

The main function of the package is called `get_chords()`. It
extracts music chords from an specific artist. There is two 
steps to obtain the data:

  1. Extraction of song urls for each music of an artist with `get_songs`.
  2. Extraction of music chords using the urls with `get_chords`.

```{r, eval =  FALSE}
library(tidyverse) 
set.seed(20191)


# Step 1: Getting the chords for some Janis Joplin songs
songs <- "janis-joplin" %>% 
  chorrrds::get_songs() %>% 
  dplyr::sample_n(5)        # Just selecting a random sample of 5 songs 

# Step 2: getting the chords for the selected songs
chords <- songs %>% 
  dplyr::pull(url) %>%                     
  purrr::map(chorrrds::get_chords) %>%     # Mapping the function over the 
                                           # selected urls
  purrr::map_dfr(dplyr::mutate_if, is.factor, as.character)   %>% 
  chorrrds::clean(message = FALSE)         # Cleans the dataset, in case
                                           # strange elements, as music lyrics, 
                                           # are present when they shouldn't

chords %>% slice(1:10) 
```

```{r, warning=FALSE, results='asis', echo = FALSE}
library(tidyverse) 
set.seed(20191)


# Step 1: Getting the chords for some Janis Joplin songs
songs <- "janis-joplin" %>% 
  chorrrds::get_songs() %>% 
  dplyr::sample_n(5)        # Selecting a random sample of 5 pieces 

# Step 2: getting the chords for the selected songs
chords <- songs %>% 
  dplyr::pull(url) %>%                     
  purrr::map(chorrrds::get_chords) %>%     # Mapping the function over the 
                                           # selected urls
  purrr::map_dfr(dplyr::mutate_if, is.factor, as.character)   %>% 
  chorrrds::clean(message = FALSE)         # Cleans the dataset, in case
                                           # strange elements, as music lyrics, 
                                           # are present when they shouldn't

chords %>% slice(1:10) %>% knitr::kable()
```


The table above shows us how are the results of the `get_chords`
function. As you can see, the data is in a long format: the chords 
appear in the sequence they are in each song, being repeated
sometimes. 


## Data

There are actually many datasets that come built-in
with the package. They have been previously used in my
undergrad thesis, so they were just kept in the package to
work as example data. These data are composed of several 
Brazilian artists music chords. You can check the available 
datasets with the  code above: 

```{r, eval = FALSE}
library(chorrrds)
ls("package:chorrrds")                 
```


## Use case

Returning to the data we collected before, let's explore it! 

The first thing we can look at is the most common chords in 
each music. Which are the common chords in music made
by Janis Joplin? Are the proportions of these chords
similar between the songs? 

```{r, eval = FALSE}
chords %>% 
  dplyr::group_by(song) %>% 
  dplyr::count(chord) %>%
  dplyr::top_n(n, n = 3) %>%
  dplyr::mutate(prop = scales::percent(n/sum(n))) 
```


```{r, echo = FALSE}
chords %>% 
  dplyr::group_by(song) %>% 
  dplyr::count(chord) %>%
  dplyr::top_n(n, n = 3) %>%
  dplyr::mutate(prop = scales::percent(n/sum(n))) %>% 
  knitr::kable()
```

With the dataset analyzed here, we can already
obtain some interesting information. For instance, 
we can observe a quite varying difference between the 
most common chords proportions, there is no clear rule. 
That shows us that the structure of her songs 
might not follow a closed pattern, which can be a sign of how 
creative the artist was. 

We can also look at something called "chord bigrams". This
is pretty much the task of creating pairs of chords that 
happened in sequence, by song, and analyze their frequencies, 
and it's done with the `chords_ngram()` function

```{r, eval = FALSE}
chords %>%
  split(.$song) %>% 
  purrr:::map(chorrrds::chords_ngram, n = 2) %>% 
  dplyr::bind_rows() %>% 
  dplyr::group_by(song) %>% 
  dplyr::count(chords_ngram) %>% 
  dplyr::top_n(n, n = 2) 
```

```{r, echo = FALSE}
chords %>%
  split(.$song) %>% 
  purrr:::map(chorrrds::chords_ngram, n = 2) %>% 
  dplyr::bind_rows() %>% 
  dplyr::group_by(song) %>% 
  dplyr::count(chords_ngram) %>% 
  dplyr::top_n(n, n = 2) %>% 
  knitr::kable()
```

There are some bigrams that happen a lot in a song, while 
others just a few times, but are still the most frequent ones. In 
the song called "A Woman Left Lonely", for example,
we have only two chords sequences(Bb -> F and G -> C), 
which is harmonically interesting to observe. 


Now, we have already explored the data a little bit. We can
make it even more interesting, by building a chord diagram. 
The word "chord" here does not mean the musical one, but a
graphic element that shows us the strength of a connection. 
The connections will be the observed chord transitions in our
selected songs, and their strengths, how many times each
transition happened. With this configuration, the chord
diagram makes the relationship between each chord
explicit. 

```{r, warning = FALSE}
# devtools::install_github("mattflor/chorddiag")
library(chorddiag)

comp <- chords %>% 
  dplyr::mutate(seq = lead(chord)) %>% 
  dplyr::filter(chord != seq) %>% 
  dplyr::group_by(chord, seq) %>%  
  dplyr::summarise(n = n())

mat <- tidyr::spread(comp, key = chord, value = n, fill = 0)  
mm <- as.matrix(mat[, -1]) 

# Building the chords diagram
chorddiag::chorddiag(mm, showTicks = FALSE,
                     palette = "Reds")
```

Now we can clearly see how the transitions behave 
in the songs we're using. There are strong 
connections between the chord A and E, A and 
B, and the others one are, in general, fragmented.  
A cool thing to notice is that the diagram
is interactive, then we can see the strength 
of each transition with the mouse cursor!  

## Wrap up

In this blog post, we:   

  - introduced the `chorrrds` package, which
  extracts chords from the songs of an artist.
  - we used Janis Joplin songs as an example of how
  the extraction is done. 
  - we presented a brief analysis of the
  music chords, along with a chord diagram.  
  
`chorrrds` is a new package, with a lot of 
potential and many possible applications
to be explored. We hope this was useful, and 
that now you're starting to be as enchanted by
music information retrieval as we are!



