---
title: "Introduction to the chorrrds package"
description: |
  Music chords analysis!
author:
  - name: "Bruna Wundervald"
    url: https://brunaw.com
    affiliation: Maynooth University
    affiliation_url: https://www.maynoothuniversity.ie/
  - name: "Julio Trecenti"
    url: https://github.com/jtrecenti
    affiliation: Curso-R
    affiliation_url: http://curso-r.com/
date: 08-19-2018
output:
  radix::radix_article:
    self_contained: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro

`chorrrds` is a package to retrieve and analyse music data. It scrapes 
the [*Cifraclub*]() website to download and organize music chords.

The main reason to create `chorrrds` was Bruna's undergrad thesis. In 
her work, she did an end-to-end analysis, exploring feature engineering
techniques to describe and predict musical genres from music chord 
representation.

`chorrrds` can be considered a package for MIR (Music Information 
Retrieval). MIR is a broad area of computational music which extracts 
and processes music data, from the unstructured ones, as sound waves,
to structured, as sheet music or chords.

In this post we'll describe `chorrrds` functions and show 
some examples. Stay tuned!

## Installation

You can install `chorrrds` from your favourite CRAN mirror, simply running:

```{r, eval = FALSE}
install.packages("chorrrds")
```

You can also install the latest versios of `chorrrds` from the 
R-Music GitHub organization with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("r-music/chorrrds")
```

## Functions

The main function of the package is called `get_chords()`. It
extracts music chords from an specific artist. There is two 
steps to obtain the data:

  1. Extraction of song urls for each music of an artist with `get_songs`.
  2. Extraction of music chords using the urls with `get_chords`.

```{r, warning=FALSE, results='asis'}
library(tidyverse) 
set.seed(20191)


# Step 1: Getting the chords for some Janis Joplin songs
songs <- "janis-joplin" %>% 
  chorrrds::get_songs() %>% 
  dplyr::sample_n(5)        # Selecting a random sample of 5 pieces 

# Step 2: getting the chords for the selected songs
chords <- songs %>% 
  dplyr::pull(url) %>%                     
  purrr::map(chorrrds::get_chords) %>%     # Mapping the function over the 
                                           # selected urls
  purrr::map_dfr(dplyr::mutate_if, is.factor, as.character)   %>% 
  chorrrds::clean(message = FALSE)         # Cleans the dataset, in case
                                           # strange elements, as music lyrics, 
                                           # are present when they shouldn't

chords %>% slice(1:10) %>% knitr::kable()
```

The table above shows us how are the results of the `get_chords`
function. As you can see, the data is in a long format: the chords 
appear in the sequence they are in each music, being repeated
sometimes. The `music` column contains the name of the artist
and the name of the song. This can be changed if preferred, 
with: 

```{r}
chords <- chords %>% 
  tidyr::separate(music, c("artist", "music"), 
                  sep = "(?<=joplin) ", extra = "merge")


chords %>% slice(1:10) %>% knitr::kable()
```


## Data

There are many datasets that come with the package. They were
used in the undergrad thesis, so it was a choice just to keep it 
in the package. The data is composed of several Brazilian artists
music chords. You can check the available datasets with the 
code above, which won't be run here because the results 
are long: 

```{r, eval = FALSE}
library(chorrrds)
ls("package:chorrrds")                 
```


## Use case

Returning to the data we collected before, let's explore it! 

The first thing we can look at is the most common chords in 
each music. Which are the common chords in music made
by Janis Joplin? Are the proportions of these chords
similar between the songs? 

```{r}
chords %>% 
  dplyr::group_by(music) %>% 
  dplyr::count(chord) %>%
  dplyr::top_n(n, n = 3) %>%
  dplyr::mutate(prop = scales::percent(n/sum(n))) %>% 
  knitr::kable()
```

With the dataset analyzed here, we can already
obtain some interesting information. For some of it, 
as the first and second pieces, the 3 most common chords 
appeared in a close proportion. For the others,
this happens in a different way. Both the proportions and
the absolute quantities of the chords vary more.
That shows us that the structure of her songs 
don't follow a closed pattern, which can be a sign of how 
creative the artist was. 

We can also look at something called "chord bigrams". This
is pretty much the task of creating pairs of chords that 
happened in sequence, by music, and analyze their frequencies. 

```{r}
chords %>%
  dplyr::group_by(music) %>% 
  tidytext::unnest_tokens(bigram, chord, to_lower = FALSE,
                          token = "ngrams", n = 2) %>% 
  dplyr::count(bigram) %>% 
  dplyr::top_n(n, n = 2) %>%
  knitr::kable()
```

There are some bigrams that happen a lot in a song, while 
others just a few times, but are still the most frequent ones. In 
the song called "Piece of my heart", we have the repetition
of the chord "B", which is described by the appearance of 
the "B B" bigram. 


Now, we have already explored the data a little bit. We can
make it even more interesting, by building a chord diagram. 
The word "chord" here does not mean the musical one, but a
graphic element that shows us the strength of a connection. 
The connections will be the observed chord transitions in our
selected songs, and their strengths, how many times each
transition happened. With this configuration, the chord
diagram makes the relationship between each chord
explicit. 

```{r, warning = FALSE}
# devtools::install_github("mattflor/chorddiag")
library(chorddiag)

comp <- chords %>% 
  dplyr::mutate(seq = lead(chord)) %>% 
  dplyr::filter(chord != seq) %>% 
  dplyr::group_by(chord, seq) %>%  
  dplyr::summarise(n = n())

mat <- tidyr::spread(comp, key = chord, value = n, fill = 0)  
mm <- as.matrix(mat[-19, -1]) 

# Building the chords diagram
chorddiag::chorddiag(mm, showTicks = FALSE,
                     palette = "Reds")
```

Now we can clearly see how the transitions behave 
in the songs we're using. There are strong 
connections between the chord A and E, A and 
B, and the others one are, in general, fragmented.  
A cool thing to notice is that the diagram
is interactive, then we can see the strength 
of each transition with the mouse cursor!  

## Wrap up

In this blog post, we:   

  - introduced the `chorrrds` package, which
  extracts chords from the songs of an artist.
  - we used Janis Joplin songs as an example of how
  the extraction is done. 
  - we presented a brief analysis of the
  music chords, along with a chord diagram.  
  
`chorrrds` is a new package, with a lot of 
potential and many possible applications
to be explored. We hope this was useful, and 
that now you're starting to be as enchanted by
music information retrieval as we are!



