---
title: "Connecting packages"
description: |
  R package to connect with the Spotify API.
preview: https://github.com/r-music/site/blob/master/img/Rspotify/logo.png?raw=true
author:
  - name: "Bruna Wundervald"
    url: https://brunaw.com
    affiliation: Maynooth University
    affiliation_url: https://www.maynoothuniversity.ie/
date: 03-13-2019
bibliography: bibli.bib
output:
  radix::radix_article:
    self_contained: false
editor_options: 
  chunk_output_type: console
---  


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Today we are going to talk a little about how to 'connect' the data
obtained from different R-Music packages. For example, 
the `chorrds`, `vagalumeR` and `Rspotify` packages can extract data for
all of the songs of an artist. However, sometimes the name of the songs
might not be written the same in the APIs or website
(in the case of the `chorrrds` package). This means that some extra work
to correct these disagreements between the names might be necessary,
if we don't want to lose too many observations when joining data
from different sources. 

This situation is actually very common in a diversity of contexts. 
Very often, we need to fix strings and joining keys in our data, 
specially when it comes from public sources. Because of that, the topic
of this post is not restricted to the application to music data. 


First, we need to load the necessary packages for obtaining the data 
(and `tidyverse`, of course). A detailed
explanation of how to do so for each package considered
is presented [here](https://r-music.rbind.io/posts/2018-08-19-chords-analysis-with-the-chorrrds-package/),
[here](https://r-music.rbind.io/posts/2018-11-22-introduction-to-the-vagalumer-package/)
and [here](https://r-music.rbind.io/posts/2018-10-01-rspotify/).  

```{r}
library(tidyverse)
library(chorrrds)
library(vagalumeR)
library(Rspotify)

# Setting our artist of interest
artist <- "muse"
```

Here, we'll be working with the data about *Muse*,  
a very well known British rock band. Below, we proceed by obtaining the
music chords for the band: 

```{r, eval = FALSE}
# Getting the chords
songs_chords <- get_songs(artist) 
chords <- get_chords(songs_chords$url)
```

```{r, echo = FALSE}
chords <- read.table("data/muse/chords_muse.txt")
```

```{r}
chords %>% glimpse()
``` 

From the `glimpse` of the data, we can see that everything is as 
expected. Next, we obtain the lyrics: 

```{r}
lyrics <- read.table("data/muse/lyrics_muse.txt")
```

```{r, eval = FALSE}
key_vg <- "your-api-key-to-vagalume"
# song names retrieval to get the lyrics
song <- songNames(artist)

lyrics <- song %>% 
  dplyr::pull(song.id) %>% 
  purrr::map_df(lyrics, 
                artist = artist,
                type = "id", 
                key = key_vg)
```

```{r}
lyrics %>% glimpse()
``` 

All good here too. The `Spotify` data is extracted with: 

```{r, eval = FALSE}
key_sp <- "your-authentication-key-to-spotify"

find_artist <- searchArtist(artist, token = key_sp)
muse_albums <- getAlbums(find_artist$id[1], token = key_sp)

muse_songs <- muse_albums %>% 
  dplyr::pull(id) %>% 
  purrr::map_df(
    ~{
      getAlbum(.x, token = key_sp) %>% 
        dplyr::select(id, name) 
    }) %>% 
  dplyr::unnest()

features <- muse_songs %>% 
  dplyr::pull(id) %>% 
  purrr::map_df(getFeatures, token = key_sp) %>% 
  dplyr::left_join(muse_songs, by = "id")
```

```{r}
features <- read.table("data/muse/features_muse.txt")
```

```{r}
features %>% glimpse() 
```

Good! Now we have the 3 datasets, which all came from different 
sources. Our next step is to try to make the column that has the
name of the song the most similar in the tree `data.frame`, with: 

```{r}
# Adjusting the name column for our datasets
chords <- chords %>% 
  dplyr::mutate(song = stringr::str_remove(music, "muse ")) %>% 
  dplyr::select(-music)

lyrics <- lyrics %>% 
  dplyr::mutate(song = stringr::str_to_lower(song))

features <- features %>% 
  dplyr::mutate(song = stringr::str_to_lower(name)) %>% 
  dplyr::select(-name)
```

We performed just some small changes, as making the name of the songs
lower case and removing the name of the band from the song column. 
We also created a column
with the same name in the 3 datasets, so this will be standardized from
now on. When we `inner_join` the data, there are some rows that don't
match. This demonstrates that we'll have some work to do here to 
minimize the damage: 

```{r}
unite_muse <- chords %>% 
  dplyr::inner_join(lyrics, by = "song") %>% 
  dplyr::inner_join(features, by = "song")
  
nrow(chords) - nrow(unite_muse)
```

727 is the total of rows that did not match between the three datasets. 
Our next approach is to look closer in each data and think about
how we can fix it. Let's see which song the chords and the lyrics 
files *don't* have in common: 

```{r}
# Songs that didn't match between the chords and lyrics data
anti_chords_lyrics <- chords %>% 
  dplyr::anti_join(lyrics, by = "song") 

anti_chords_lyrics %>% dplyr::distinct(song)
```

We have our mismatches. First, we'll remove those
numbers from the `song` column in the chords data, and
fix two songs that we know by eye that are wrong: 

```{r}
chords <- chords %>% 
  dplyr::mutate(
    song = stringr::str_remove(song, "[0-9]{2,7}$"), 
    song = stringr::str_replace(song, " $", ""), 
    song = ifelse(
      stringr::str_detect(
        song, "map of head|map of your head"), 
      "map of your head", song),
    song = ifelse(
      stringr::str_detect(
        song, "hyper music|hyper chondriac music"), 
      "hyper music", song))
```

But we can't actually be fixing every song by eye. This can be 
exhaustive and potentially lead to mistakes. We'll use a distance measure
between the strings to get the next results. 

The `RecordLinkage` package contains the `levenshteinSim` 
function, that calculates a similarity measure based on
the Levenshtein distance for two strings. There are different 
functions in `R` that use Levenshtein, but this one in specific
has a easy interpretation, since it returns a normalized number
(between 0 and 1), being 1 the most similar possible. As for the 
Levenshtein distance, this value is given by the minimum of 
single-character edits needed to go from one string to another. 
See more in (@Yujian2007). 

Let's get back to the code. Below, we show how to find the most
similar song names of the ones that didn't exactly match:  

```{r}
library(RecordLinkage)

# Updating the anti_chords_lyrics
anti_chords_lyrics <- chords %>% 
  dplyr::anti_join(lyrics, by = "song") 

# Saving names that didn't match
names_to_fix <- anti_chords_lyrics %>% 
  dplyr::distinct(song) %>% 
  dplyr::pull(song)

# Finding the distances between each song in the lyrics data and 
# the ones that didn't match
dists <- lyrics$song %>% 
  purrr::map(levenshteinSim, str1 = names_to_fix)

# Finding the biggest similarities between the songs 
# in the lyrics data and the ones that didn't match
ordered_dists <- dists %>% purrr::map_dbl(max)
max_dists <- dists %>% purrr::map_dbl(which.max)

# Filtering to have only similarities > 0.50
indexes_min_dist <- which(ordered_dists > 0.50)
songs_min_dist <- lyrics$song[indexes_min_dist]
index_lyrics <- max_dists[which(ordered_dists > 0.50)]

# Saving the most similar string in the chords and 
# lyrics data 
results_dist_lyrics <- data.frame(
  from_chords = names_to_fix[index_lyrics],
           from_lyrics = songs_min_dist)
```

```{r, echo = FALSE}
results_dist_lyrics %>% knitr::kable()
```

The strings found with the `>0.5` criteria
are actually quite similar, right? We can now transform the similar
strings in only one version, so it will match in both datasets. We do
it with the `case_when` function, as written below: 


```{r}
lyrics <- lyrics %>% 
  dplyr::mutate(
    song = 
      dplyr::case_when( 
        song == "hate this & i'll love you" ~ "hate this and ill love you",
        song == "please, please, please let me get what i want" ~ "please please please let me get what want",          
        song == "soldier's poem" ~ "soldiers poem",
        song == "con-science" ~ "conscience",
        song == "can't take my eyes off you" ~ "cant take my eyes off you",
        song == "hyper chondriac music" ~ "hyper music",
        song ==  "i belong to you/mon coeur s'ouvre à ta voix" ~ "i belong to you mon coeur souvre toi",
        song ==  "neutron star collision (love is forever)" ~ "neutron star collision",
        TRUE ~ song))

```

Now the `anti_join` of the two datasets is smaller and we are losing
way less information :) 

```{r}
chords %>%  dplyr::anti_join(lyrics, by = "song") %>% dplyr::distinct(song) 
```


Note that you can use `paste` to create the syntax for `case_when` 
automatically, something like:

```{r, eval = FALSE}
paste0(
  paste0("song == ", "'", results_dist_lyrics$from_chords, "' ~ '", 
         results_dist_lyrics$from_lyrics, "', "), collapse = "")
```

This was for the chords and lyrics datasets. We are setting the 
chords data as the 'baseline', so every string we correct, is being
corrected to the version found in the chords data. In the following, 
we do the same procedure for the features data, we try to make
it closer to the song names available in the chords extraction: 

```{r}
# Songs that didn't match between the chords and features data
anti_chords_features <- chords %>% 
  dplyr::anti_join(features, by = "song") 

anti_chords_features %>% dplyr::distinct(song)
```

And finding the very similar strings: 

```{r}
names_to_fix <- anti_chords_features %>% 
  dplyr::distinct(song) %>% 
  dplyr::pull(song)

dists <- features$song %>% 
  purrr::map(levenshteinSim, str1 = names_to_fix)

ordered_dists <- dists %>% purrr::map_dbl(max)
index_dists <- dists %>% purrr::map_dbl(which.max)

indexes_min_dist <- which(ordered_dists > 0.50)
songs_min_dist <- features$song[indexes_min_dist]
index_features <- index_dists[which(ordered_dists > 0.50)]

results_dist_features <- data.frame(
  from_chords = names_to_fix[index_features],
  from_features = songs_min_dist)
```

```{r}
results_dist_features %>% knitr::kable()
```

Again, we have strings that are very close and probably refer to the 
same song (we can see it by eye). Let's use this information
to correct the names of the song in the lyrics data: 

```{r}
features <- features %>% 
  dplyr::mutate(
    song = dplyr::case_when( 
      song == "united states of eurasia (+collateral damage)" ~ 
        "united states of eurasia",
      song == "i belong to you (+mon coeur s'ouvre a ta voix)" ~ "i belong to you mon coeur souvre toi",          
      song == "soldier's poem" ~ "soldiers poem",
      song == "darkshines" ~ "dark shines",
      song == "hate this and i'll love you" ~ "hate this and ill love you",
      TRUE ~ song))

chords %>%   
  dplyr::anti_join(features, by = "song") %>% 
  dplyr::distinct(song)
```

We reduced a little bit more our mismatches in the three datasets. 
What can also happen is that a mismatch is not a result from 
different ways of writing a song name, but of songs that really don't
exist in all the three datasets. The final rows
difference between the original and final data is: 

```{r}
final_data <- chords %>% 
  dplyr::inner_join(lyrics, by = "song") %>% 
  dplyr::inner_join(features, by = "song") 

nrow(chords) - nrow(final_data) 
```

So we managed to 'save' more than 300 rows of the data by doing some small
fixes in the key that connects the three datasets, and there is still
plenty of room for improvements. The final glimpse of the data is:

```{r}
final_data %>% glimpse() 
```

Our final dataset combines 3 data sources and is very rich. We won't
analyze it at this moment, but it's clear that the possibilities
are infinite here. 

## Wrap-up

To combine datasets from different sources, one often challenge is:
fix the mismatchings in the joining key. Here, we joined data from
the `chorrds`, `vagalumeR` and `Rspotify` packages. The approaches
to deal with mismatching song names were:

  - Standardized column names and lower cases; 
  - Fix some cases by eye;
  - Use a distance measure to find the most similar strings. 

We ended up by saving more than 300 rows of observations that
would be wasted otherwise, and finished with a complete dataset,
that contains information about the chords, the lyrics and the 
Spotify variables of the *Muse* songs.  

